---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  message = FALSE, 
  warning = FALSE,
  cache = TRUE
)
```

# About

[![Travis-CI Build Status](https://travis-ci.org/jjchern/ahrf.svg?branch=master)](https://travis-ci.org/jjchern/ahrf)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/jjchern/ahrf?branch=master&svg=true)](https://ci.appveyor.com/project/jjchern/ahrf)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ahrf)](https://cran.r-project.org/package=ahrf)

This repo contains R scripts (in the [`data-raw` folder](https://github.com/jjchern/ahrf/tree/master/data-raw)) that download county-level and state-level 
[Area Health Resources Files (AHRF)](http://ahrf.hrsa.gov/download.htm). The datasets are stored in the [`data` folder](https://github.com/jjchern/ahrf/tree/master/data). 

AHRF is issued annually. 
The most recent release is in 2016 (as of July 24, 2017).

For more information on the AHRF files, see [https://datawarehouse.hrsa.gov/topics/ahrf.aspx](https://datawarehouse.hrsa.gov/topics/ahrf.aspx).

# Installation

You can also download the datasets as an R package.
The size of `ahrf_county.rda` is 17.5M, so it might take a while to install and load into memory.

```R
# install.packages("devtools")
devtools::install_github("jjchern/ahrf")

# To uninstall the package, use:
# remove.packages("ahrf")
```

# Usage

## Load the state file

```{r}
library(tidyverse)
ahrf::ahrf_state
```

## There're 3230 rows and 6921 columns in the county file (wide format)

```{r}
dim(ahrf::ahrf_county)
```

## Variable labels are included

```{r}
library(labelled)
ahrf::ahrf_county %>% 
        select(F04437, F00002, contains("F08921"), contains("F11984")) %>% 
        var_label() %>% 
        enframe() %>% 
        unnest()
```

## County-level hospital beds in 2013

```{r}
ahrf::ahrf_county %>% 
        select(county = F04437, 
               fips = F00002, 
               beds_2013 = `F08921-13`,
               pop_2013 = `F11984-13`) %>% 
        mutate(beds_2013 = as.integer(beds_2013),
               pop_2013 = as.integer(pop_2013),
               beds_2013_p10k = beds_2013 / pop_2013 * 10000) -> beds
beds

lapply(beds, summary)
```
